Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-25 paper.tex
Minor updates

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, in a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g. RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although it
might as well be delivered by the workflow engine directly to the
user.

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the system
in~\cite{wu2010accelerating} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g. pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The architectures described in Section~\ref{sec:architectures} are
evaluated in Table~\ref{table:evaluation}. We use 5 main criteria to
evaluate the architectures: integration effort, robustness of workflow
execution, extensibility, scalability and other specific features. Criteria
break down to specific metrics where \emph{low value indicates better
performance}. For each criterion, a total score is computed by
summing up the individual metrics. We ensure that the different
metrics in a criterion measure similar entities so that summation
does not incur in loss of accuracy. The
criteria and metrics are explained hereafter.

@@

\emph{Extensibility} measures the difficulty to replace or add
elements in the architecture. It is determined as the number of
interactions and components that need to be modified when a new
element is added. Modification of a component is required when its
code needs to be updated or recompiled (science gateway or workflow
service), or when a new piece of software has to be installed
(infrastructure only). Modification of an interaction is deemed
necessary when the parameters involved in this interaction are
modified.  Extensibility breaks down in 4 metrics depending on the
type of element that has to be added or replaced:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item New engine type (\texttt{E$_1$}): number of interactions or
  components to modify to integrate a new type of workflow engine in
  the architecture. Workflow engines belong to different types when
  they cannot be invoked using the same interface. Adding a new type
  of workflow engine allows to execute more workflows in the science
  gateway.
\item New engine version (\texttt{E$_2$}): number of interactions or
  components to modify to integrate a new version of a workflow engine
  in the architecture, assuming that another version of the same
  engine type is already available. Different versions of a workflow
  engine share the same interface, i.e. they can be invoked using the
  same software. When this is not the case, the different versions are
  considered as different engine types.
\item New workflow (\texttt{E$_3$}): adding a new workflow is a very
  common operation that does not require modifying software components
  or interactions. We measure the difficulty to integrate a new
  workflow by counting the number of interactions required
  to integrate a new workflow in the architecture, assuming
  that the engine type and version required to execute this workflow
  are already available.
\item New infrastructure (\texttt{E$_4$}): number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture. Adding a new infrastructure allows to provide more
  computing or storage power, to access specific types of resources
  (e.g. GPUs, clouds), or to enforce execution policies 
  (e.g. constrain data to remain in a particular network domain).\note{Rafael}{Adding a new engine type may also allow run in different infrastructures. Maybe it is worth to put a sentence to clarify that the metric account for it separately.}
%\item New science gateway: \todo{Not sure if this makes sense}.
\end{itemize}
The 4 metrics are summed to obtain a global index that measures the
difficulty to extend the architecture.

>>>>

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, in a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although it
might as well be delivered by the workflow engine directly to the
user.

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the system
in~\cite{wu2010accelerating} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g., pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The architectures described in Section~\ref{sec:architectures} are
evaluated in Table~\ref{table:evaluation}. We use 5 main criteria to
evaluate the architectures: integration effort, robustness of workflow
execution, extensibility, scalability and other specific
features. Criteria break down to specific metrics where \emph{lower
  value indicates better performance}. For each criterion, a total
score is computed by summing up the individual metrics. We ensure that
the different metrics in a criterion measure similar entities so that
they can be summed. The criteria and metrics are explained hereafter.

@@

\emph{Extensibility} measures the difficulty to replace or add
elements in the architecture. It is determined as the number of
interactions and components that need to be modified when a new
element is added. Modification of a component is required when its
code needs to be updated or recompiled (science gateway or workflow
service), or when a new piece of software has to be installed
(infrastructure only). Modification of an interaction is deemed
necessary when the parameters involved in this interaction are
modified.  Extensibility breaks down in 4 metrics depending on the
type of element that has to be added or replaced:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item New engine type (\texttt{E$_1$}): number of interactions or
  components to modify to integrate a new type of workflow engine in
  the architecture. Workflow engines belong to different types when
  they cannot be invoked using the same interface. Adding a new type
  of workflow engine allows to execute more workflows in the science
  gateway.
\item New engine version (\texttt{E$_2$}): number of interactions or
  components to modify to integrate a new version of a workflow engine
  in the architecture, assuming that another version of the same
  engine type is already available. Different versions of a workflow
  engine share the same interface, i.e. they can be invoked using the
  same software. When this is not the case, the different versions are
  considered as different engine types.
\item New workflow (\texttt{E$_3$}): adding a new workflow is a very
  common operation that does not require modifying software components
  or interactions. We measure the difficulty to integrate a new
  workflow by counting the number of interactions required
  to integrate a new workflow in the architecture, assuming
  that the engine type and version required to execute this workflow
  are already available.
\item New infrastructure (\texttt{E$_4$}): number of interactions or components to
  modify to integrate a new type of infrastructure in the
  architecture. Adding a new infrastructure allows to provide more
  computing or storage power, to access specific types of resources
  (e.g., GPUs, clouds), or to enforce execution policies 
  (e.g., constrain data to remain in a particular network domain).\note{Rafael}{Adding a new engine type may also allow run in different infrastructures. Maybe it is worth to put a sentence to clarify that the metric account for it separately.}
%\item New science gateway: \todo{Not sure if this makes sense}.
\end{itemize}
The 4 metrics are summed to obtain a global index that measures the
difficulty to extend the architecture.
